---
title: "Technical Appendix - US Public Library Survey"
author: "Schyuler Lujan"
date: "3/24/2020"
output: html_document
---

# Project Overview

U.S. Public libraries are a crucial resource for many communities, providing access to information, services, technology, materials, programming, space for community events. Libraries serve multi-faceted roles in supporting their communities, and examining library usage reveals how different communities access (or are not accessing) this public resource. Additionally, with the growing use of new technologies and shifting public needs, libraries are faced with a changing landscape of modern library usage. Analyzing library usage and usage trends over time can indicate where these changes are most pronounced and potentially how libraries can best respond to these changees in alignment with their particular community needs.


This project will examine and produce the following:

- A decade of usage trends for U.S. Public libraries from 2008 - 2017
- Library usage benchmarking analysis for 2017

## Description of Data

For this project, I will use the Public Libraries Survey from the Institute of Museum and Library Services, for the survey years 2008 - 2017. This survey data is administered to public libraries across the United States, the District of Columbia, and U.S. territories. 

Data Source: https://www.imls.gov/research-evaluation/data-collection/public-libraries-survey

### Caveats

A few important caveats to note:

- Data points for some libraries are suppressed to protect confidentiality (ie: salary information for libraries with small staff numbers), resulting in incomplete information for some variables
- Not all libraries reported answers for every question.
- Library usage variables are highly correlated with one another, which makes it difficult to estimate the impact/effect associated with specific variables

# Data Tidying

## Load Libraries

```{r echo = FALSE, warning = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 
# Clear environment of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)
```

```{r message = FALSE, warning = FALSE}
# Load relevant libraries
library(tidyverse)
library(stringr)
library(janitor)
#library(reshape)
# Set working directory
#setwd('/Users/Tiny Moo Cow/OneDrive - Seattle University/OMSBA/Other Projects/US-Public-Libraries')
```

## Load Data
```{r message = FALSE, warning = FALSE}
# Load all datasets from 2007 thru 2017
#PLS2007 <- read.csv("./pls-data/PLS_2007.csv")
PLS2008 <- read.csv("./pls-data/PLS_2008.csv")
PLS2009 <- read.csv("./pls-data/PLS_2009.csv")
PLS2010 <- read.csv("./pls-data/PLS_2010.csv")
PLS2011 <- read.csv("./pls-data/PLS_2011.csv")
PLS2012 <- read.csv("./pls-data/PLS_2012.csv")
PLS2013 <- read.csv("./pls-data/PLS_2013.csv")
PLS2014 <- read.csv("./pls-data/PLS_2014.csv")
PLS2015 <- read.csv("./pls-data/PLS_2015.csv")
PLS2016 <- read.csv("./pls-data/PLS_2016.csv")
PLS2017 <- read.csv("./pls-data/PLS_2017.csv")
# Add SRVY_YEAR variable to each dataset to track year of survey
#PLS2007 <- mutate(PLS2007, SRVY_YEAR = 2007)
PLS2008 <- mutate(PLS2008, SRVY_YEAR = 2008)
PLS2009 <- mutate(PLS2009, SRVY_YEAR = 2009)
PLS2010 <- mutate(PLS2010, SRVY_YEAR = 2010)
PLS2011 <- mutate(PLS2011, SRVY_YEAR = 2011)
PLS2012 <- mutate(PLS2012, SRVY_YEAR = 2012)
PLS2013 <- mutate(PLS2013, SRVY_YEAR = 2013)
PLS2014 <- mutate(PLS2014, SRVY_YEAR = 2014)
PLS2015 <- mutate(PLS2015, SRVY_YEAR = 2015)
PLS2016 <- mutate(PLS2016, SRVY_YEAR = 2016)
PLS2017 <- mutate(PLS2017, SRVY_YEAR = 2017)
# Remove variables beginning with "F_" from all dataframes
#PLS2007 <- select(PLS2007, -starts_with("F_"))
PLS2008 <- select(PLS2008, -starts_with("F_"))
PLS2009 <- select(PLS2009, -starts_with("F_"))
PLS2010 <- select(PLS2010, -starts_with("F_"))
PLS2011 <- select(PLS2011, -starts_with("F_"))
PLS2012 <- select(PLS2012, -starts_with("F_"))
PLS2013 <- select(PLS2013, -starts_with("F_"))
PLS2014 <- select(PLS2014, -starts_with("F_"))
PLS2015 <- select(PLS2015, -starts_with("F_"))
PLS2016 <- select(PLS2016, -starts_with("F_"))
PLS2017 <- select(PLS2017, -starts_with("F_"))
```

## Data Merging

```{r}
# Vertical merge of PLS2008 and PLS2009------------------------------------------------------------------------
# Identify variable differences between dfs
compare_df_cols(PLS2008, PLS2009, return = c("mismatch"), bind_method = c("rbind"))
# Remove discontinuted cols from PLS2008
PLS2008 <- select(PLS2008, -CNTYFIPS, -PUB_FIPS)
# Add new cols and change col type in PLS2008
PLS2008 <- mutate(PLS2008,
                  YAATTEN = as.integer(NA),
                  YAPRO = as.integer(NA)
                  )
# Vertical merge
PLS <- rbind(PLS2008, PLS2009)
```

```{r}
# Vertical merge of PLS and PLS2010----------------------------------------------------------------------------
# Identify variable differences between dfs
compare_df_cols(PLS, PLS2010, return = c("mismatch"), bind_method = c("rbind"))
# Rename AUDIO and VIDEO variables in PLS - per the user guide definitions, the variables AUDIO and VIDEO as defined from 2007 - 2009 are the same as the variables AUDIO_PH and VIDEO_PH introduced in 2010
PLS <- rename(PLS, 
              AUDIO_PH = AUDIO, 
              VIDEO_PH = VIDEO)
# Add new variables
PLS <- mutate(PLS,
              AUDIO_DL = as.integer(NA),
              VIDEO_DL = as.integer(NA),
              )
# Remove variables discontinued as of 2010
PLS <- select(PLS, -ESUBSCRP)
# Vertical merge
PLS <- rbind(PLS, PLS2010)
```

```{r}
# Vertical merge of PLS and PLS2011----------------------------------------------------------------------------
# Identify variable differences between dfs
compare_df_cols(PLS, PLS2011, return = c("mismatch"), bind_method = c("rbind"))
# Add new cols to PLS
PLS <- mutate(PLS,
              CBSA = as.integer(NA),
              MAT_TYPE = as.integer(NA),
              MICROF = as.factor(NA)
              )
# Change col type in PLS2011
PLS2011 <- mutate(PLS2011, PHONE = as.factor(PHONE))
# Vertical merge
PLS <- rbind(PLS, PLS2011)
```

```{r}
# Vertical merge of PLS and PLS2012----------------------------------------------------------------------------
# Identify variable differences between dfs
compare_df_cols(PLS, PLS2012, return = c("mismatch"), bind_method = c("rbind"))
# Remove discontinued cols
PLS <- select(PLS, -DB_LOC, -DB_OTH, -MAT_CENT, -MAT_TYPE, -WEB_ADDR)
# Change formats of PLS2012 cols
PLS2012 <- mutate(PLS2012, 
                  CDCODE = as.factor(NA),
                  LOCALE = as.factor(NA))
# Add new cols to PLS
PLS <- mutate(PLS,
              DB_LO_OT = as.integer(NA),
              GAL = as.factor(NA),
              GALMS = as.factor(NA),
              POSTMS = as.factor(NA))
# Vertical merge
PLS <- rbind(PLS, PLS2012)
```

```{r}
# Vertical merge of PLS and PLS2013----------------------------------------------------------------------------
# Identify variable differences between dfs
compare_df_cols(PLS, PLS2013, return = c("mismatch"), bind_method = c("rbind"))
# Change col types in PLS2013
PLS2013 <- mutate(PLS2013,
                  CDCODE = as.factor(NA),
                  LOCALE = as.factor(NA),
                  MICROF = as.factor(NA),
                  )
# Add new col to PLS
PLS <- mutate(PLS,
              ELMATCIR = as.integer(NA))
# Vertical merge
PLS <- rbind(PLS, PLS2013)
```

```{r}
# Vertical merge of PLS and PLS2014----------------------------------------------------------------------------
# Identify variable differences between dfs
compare_df_cols(PLS, PLS2014, return = c("mismatch"), bind_method = c("rbind"))
# Convert col types in PLS2014
PLS2014 <- mutate(PLS2014,
                  AUDIO_DL = as.integer(AUDIO_DL),
                  PHONE = as.factor(PHONE),
                  VIDEO_DL = as.integer(VIDEO_DL))
# Add new col to PLS
PLS <- mutate(PLS, WIFISESS = as.integer(NA))
# Vertical merge
PLS <- rbind(PLS, PLS2014)
```

```{r}
# Vertical merge of PLS and PLS2015----------------------------------------------------------------------------
# Identify variable differences between dfs
compare_df_cols(PLS, PLS2015, return = c("mismatch"), bind_method = c("rbind"))
# Remove discontinued cols in PLS
PLS <- select(PLS, -DATABASE, -DB_LO_OT, -DB_ST, -FIPSCO, -FIPSPLAC, -FIPSST, -GAL, -GALMS, -POSTMS)
# Add new cols
PLS <- mutate(PLS,
              ADDRTYPE = as.factor(NA),
              EC_LO_OT = as.integer(NA),
              EC_ST = as.integer(NA),
              ELECCOLL = as.integer(NA),
              GNISPLAC = as.factor(NA),
              INCITSCO = as.integer(NA),
              INCITSST = as.integer(NA),
              MSTATUS = as.factor(NA),
              SCORE = as.numeric(NA),
              )
# Change col types for PLS2015
PLS2015 <- mutate(PLS2015,
                  CDCODE = as.factor(CDCODE),
                  LOCALE = as.factor(LOCALE),
                  PHONE = as.factor(PHONE))
# Vertical merge
PLS <- rbind(PLS, PLS2015)
```

```{r}
# Vertical merge of PLS and PLS2016----------------------------------------------------------------------------
# Identify variable differences between dfs
compare_df_cols(PLS, PLS2016, return = c("mismatch"), bind_method = c("rbind"))
# Add new cols to PLS
PLS <- mutate(PLS,
              ELCONT = as.integer(NA),
              ELINFO = as.integer(NA),
              LOCALE_MOD = as.factor(NA),
              PHYSCIR = as.integer(NA),
              REAPLOCALE = as.integer(NA),
              REAPLOCALE_MOD = as.integer(NA),
              TOTCOLL = as.integer(NA)
              )
# Change col type for PLS2016
PLS2016 <- mutate(PLS2016,
                  CDCODE = as.factor(CDCODE),
                  PHONE = as.factor(PHONE))
# Vertical merge
PLS <- rbind(PLS, PLS2016)
```

```{r}
# Vertical merge of PLS and PLS2017----------------------------------------------------------------------------
# Identify variable differences between dfs
compare_df_cols(PLS, PLS2017, return = c("mismatch"), bind_method = c("rbind"))
# Remove discontinued cols
PLS <- select(PLS, -ADDRTYPE, -MSTATUS, -REAPLOCALE, -SCORE)
# Add new cols to PLS
PLS <- mutate(PLS,
              GEOMATCH = as.factor(NA),
              LOCALE_ADD = as.integer(NA),
              REAPLOCALE_ADD = as.integer(NA))
# Change type on cols for PLS2017
PLS2017 <- mutate(PLS2017,
                  CDCODE = as.factor(CDCODE),
                  LOCALE_MOD = as.factor(LOCALE_MOD),
                  PHONE = as.factor(PHONE),
                  LOCALE = as.factor(NA))
# Vertical merge
PLS <- rbind(PLS, PLS2017)
```

## Data Tidying
```{r}
# Create Unique Identifier
PLS <- mutate(PLS, ID = as.character(paste(FSCSKEY, SRVY_YEAR, sep = "-")))
# Check for duplicate values on unique identifier
sum(duplicated(PLS$ID))
# Replace values of -1 with NA (-1: missing)
PLS[PLS == -1] <- NA
# Replace values of -3 with NA (-3: closed or temporary closed administrative entity)
PLS[PLS == -3] <- NA
# Replace values of -9 with NA (-9: data suppressed to protect confidentiality)
PLS[PLS == -9] <- NA
```

```{r}
# Back fill values for 2008 - 2016 for LOCALE_ADD--------------------------------------------------------------
# Create new df for back-filling values
locale_df <- select(PLS, FSCSKEY, SRVY_YEAR, LOCALE_ADD) %>% 
  mutate(SRVY_YEAR = as.character(SRVY_YEAR),
         SRVY_YEAR = paste0("Y_", SRVY_YEAR)) %>% #to avoid issues with col names
  pivot_wider(names_from = SRVY_YEAR, values_from = LOCALE_ADD)
# Fill values from 2008 - 2016 from 2017 values (based on lat long)
locale_df <- mutate(locale_df,
                    Y_2008 = Y_2017,
                    Y_2009 = Y_2017,
                    Y_2010 = Y_2017,
                    Y_2011 = Y_2017,
                    Y_2012 = Y_2017,
                    Y_2013 = Y_2017,
                    Y_2014 = Y_2017,
                    Y_2015 = Y_2017,
                    Y_2016 = Y_2017,) %>% 
  pivot_longer(2:11, names_to = "Year", values_to = "LOCALE_ADD_FULL")
# Remove "Y_" from prefix of Year values
locale_df$Year <- str_replace_all(locale_df$Year, "Y_", "")
# Create unique identifier for merging to PLS and remove unnecessary variables
locale_df <- mutate(locale_df, ID = paste0(FSCSKEY, "-", Year)) %>% 
  select(-Year, -FSCSKEY)
# Left outer join with PLS
PLS <- left_join(PLS, locale_df, by = "ID")
# Remove LOCALE_ADD and rename LOCALE_ADD_FULL to LOCALE_ADD
PLS <- select(PLS, -LOCALE_ADD)
PLS <- rename(PLS, LOCALE_ADD = LOCALE_ADD_FULL)
```

Replace factor variable numbers with text:
```{r}
# Replace OBEREG number codes with region names----------------------------------------------------------------
PLS$OBEREG[PLS$OBEREG == 1] <- "New England"
PLS$OBEREG[PLS$OBEREG == 2] <- "Mid East"
PLS$OBEREG[PLS$OBEREG == 3] <- "Great Lakes"
PLS$OBEREG[PLS$OBEREG == 4] <- "Plains"
PLS$OBEREG[PLS$OBEREG == 5] <- "Southeast"
PLS$OBEREG[PLS$OBEREG == 6] <- "Southwest"
PLS$OBEREG[PLS$OBEREG == 7] <- "Rocky Mountains"
PLS$OBEREG[PLS$OBEREG == 8] <- "Far West"
PLS$OBEREG[PLS$OBEREG == 9] <- "Outlying Areas"
# Change variable to factor
PLS <- mutate(PLS, OBEREG = as.factor(OBEREG))
# Replace LOCALE_MOD number codes with locale names------------------------------------------------------------
PLS <- mutate(PLS, LOCALE_NAME = NA)
PLS$LOCALE_NAME[PLS$LOCALE_ADD == 11] <- "City, Large"
PLS$LOCALE_NAME[PLS$LOCALE_ADD == 12] <- "City, Mid-Size"
PLS$LOCALE_NAME[PLS$LOCALE_ADD == 13] <- "City, Small"
PLS$LOCALE_NAME[PLS$LOCALE_ADD == 21] <- "Suburb, Large"
PLS$LOCALE_NAME[PLS$LOCALE_ADD == 22] <- "Suburb, Mid-Size"
PLS$LOCALE_NAME[PLS$LOCALE_ADD == 23] <- "Suburb, Small"
PLS$LOCALE_NAME[PLS$LOCALE_ADD == 31] <- "Town, Fringe"
PLS$LOCALE_NAME[PLS$LOCALE_ADD == 32] <- "Town, Distant"
PLS$LOCALE_NAME[PLS$LOCALE_ADD == 33] <- "Town, Remote"
PLS$LOCALE_NAME[PLS$LOCALE_ADD == 41] <- "Rural, Fringe"
PLS$LOCALE_NAME[PLS$LOCALE_ADD == 42] <- "Rural, Distant"
PLS$LOCALE_NAME[PLS$LOCALE_ADD == 43] <- "Rural, Remote"
# Create variable LOCALE_CSTR for high-level groupings
PLS <- mutate(PLS, LOCALE_CSTR = if_else(startsWith(LOCALE_NAME, "City"), "City",
                                         ifelse(startsWith(LOCALE_NAME, "Suburb"), "Suburb",
                                                ifelse(startsWith(LOCALE_NAME, "Town"), "Town",
                                                       ifelse(startsWith(LOCALE_NAME, "Rural"),"Rural", NA)
                                                       )
                                                )
                                         )
              )
# Convert LOCALE_NAME to factor and remove LOCALE
PLS <- mutate(PLS, 
              LOCALE_NAME = as.factor(LOCALE_NAME),
              LOCALE_CSTR = as.factor(LOCALE_CSTR)) %>% 
  select(-LOCALE_ADD)
# Replace C_LEGBAS number codes with legal basis names----------------------------------------------------------
PLS <- mutate(PLS, LEGAL_BASIS = NA)
PLS$LEGAL_BASIS[PLS$C_LEGBAS == 1] <- "City/County"
PLS$LEGAL_BASIS[PLS$C_LEGBAS == 2] <- "Municipal Government"
PLS$LEGAL_BASIS[PLS$C_LEGBAS == 3] <- "County/Parish"
PLS$LEGAL_BASIS[PLS$C_LEGBAS == 4] <- "Library District"
PLS$LEGAL_BASIS[PLS$C_LEGBAS == 5] <- "Multi-Jurisdictional"
PLS$LEGAL_BASIS[PLS$C_LEGBAS == 6] <- "Native American Tribal Government"
PLS$LEGAL_BASIS[PLS$C_LEGBAS == 7] <- "Non-Profit Association or Agency"
PLS$LEGAL_BASIS[PLS$C_LEGBAS == 8] <- "School District"
PLS$LEGAL_BASIS[PLS$C_LEGBAS == 9] <- "Other"
# # Convert LEGAL_BASIS to factor and remove C_LEGBAS
PLS <- mutate(PLS, LEGAL_BASIS = as.factor(C_LEGBAS)) %>% 
  select(-C_LEGBAS)
```

Change names to proper case:
```{r}
# Change library name, city, and county to proper case
PLS <- mutate(PLS, 
              LIBNAME = str_to_title(LIBNAME),
              CITY = str_to_title(CITY),
              CNTY = str_to_title(CNTY),
              )
```


## Variable Categories

After reviewing the 2007 and 2017 documentation and the criteria used by the Library Journal rating system, I have determined that the following variable categories will be assessed for this analysis and eventual dashboard development:

- Service outlets
- Operating revenue
- Collection expenditures
- Library collection
- Library services
- Circulation of materials
- Library programs
- Electronic technology

We are most interested in assessing how different library characteristics (ie: geographic location, population, urban status), library collection, available resources to patrons, and sources of operating revenue affect library usage. We are also interested in any trends over the 10-year period 2007 to 2017. 

### Identification

- `FSCSKEY`: Library identification code (unique)
- `LIBNAME`: Library name (for administrative entity)

### Location Characteristics

- `POPU LSA`: Population of the legal service area
- `LEGAL_BASIS`: Legal basis code (ie: city/county, municipal govt, etc.)
  - CC: City/County
  - CI: Municipal Government
  - CO: County/Parish
  - LD: Library District
  - MJ: Multi-jurisdictional
  - NL: Native American Tribal Government
  - NP: Non-Profit Association or Agency
  - SD: School District
  - OT: Other
- `LOCALE_NAME`: Urban-centric locale code based on latitude and longitude of administrative entity*
  - 11: City, Large
  - 12: City, Mid-Size
  - 13: City, Small
  - 21: Suburb, Large
  - 22: Suburb, Mid-Size
  - 23: Suburb, Small
  - 31: Town, Fringe
  - 32: Town, Distant
  - 33: Town, Remote
  - 41: Rural, Fringe
  - 42: Rural, Distant
  - 43: Rural, Remote
  
*Variable was added in 2017; since this is based off of lat long coordinates, I back-filled the values for this variable for prior years based off of 2017 values

### Geography

- `STABR`: State code, two-letter
- `CITY`: City or town (for administrative entity)
- `ZIP`: Five-digit zip code
- `CNTY`: County
- `OBEREG`: Region
  - 1: New England
  - 2: Mid East
  - 3: Great Lakes
  - 4: Plains
  - 5: Southeast
  - 6: Southwest
  - 7: Rocky Mountaints
  - 8: Far West
  - 9: Outlying Areas
- `LONGITUD`: Longitude of location
- `LATITUDE`: Latitude of location

### Service Outlets

- `CENTLIB`: Number of central libraries
- `BRANLIB`: Number of branch libraries
- `BKMOB`: Number of bookmobiles
- `LIBRARIA`: Number of FTE employees with title of librarian

### Operating Revenue

- `LOCGVT`: Operating revenue from local government
- `STGVT`: Operating revenue from state government
- `FEDGVT`: Operating revenue from federal government
- `OTHINCM`: Other operating revenue

### Operating Expenditures

- `STAFFEXP`: Total staff expenditures, `SALARIES` + `BENEFIT`
**Collection Expenditures**
- `PRMATEXP`: Operating expenditures for print materials
- `ELMATEXP`: Operating expenditures for electronic materials
- `OTHMATEX`: Operating expenditures for all other library materials
- `TOTEXPCO`: Total expenditures on library collection
**Other Operating Expenditures**
- `OTHOPEXP`: Other operating expenditures
- `TOTOPEXP`: Total operating expenditures, `STAFFEXP`+`TOTEXPCO`+`OTHOPEXP`

### Library Collection

- `BKVOL`: Print materials
- `EBOOK`: Electronic books
- `AUDIO_PH`: Audio materials (physical)
- `AUDIO_DL`: Downloadable audio units
- `VIDEO_PH`: Video materials (physical)
- `VIDEO_DL`: Downloadable video units
- `ELECCOLL`: Total electronic collections

### Library Services

- `VISITS`: Total annual library visits
- `REGBOR`: Registered borrowers

### Circulation

- `TOTCIR`: Total annual circulation transactions
- `KIDCIRCL`: Total annual circulation of all children's materials
- `ELMATCIR`: Eletronic material circulation
- `PHYSCIR`: Total annual physical material circulation
- `TOTCOLL`: Total annual count of physical item circulation, circulation of electronic material and successful retrieval of electronic information

### Inter-Library Loans
- `LOANTO`: Total annual loans provided to other libraries
- `LOANFM`: Total annual loans received from other libraries

### Library Programs
- `TOTPRO`: Total library programs
- `KIDPRO`: Total children's programs
- `YAPRO`: Total Young Adult Programs
- `TOTATTEN`: Total audience at all library programs
- `KIDATTEN`: Total audience at all children's programs
- `YAATTEN`: Young adult program attendance

### Electronic Technology
- `GPTERMS`: Internet terminals used by general public
- `PITUSR`: Users of public internet computers per year
- `WIFISESS`: Total annual wireless sessions provided by the library


Select Variables:
```{r}
# Select variables for final analysis
PLS_FINAL <- select(PLS, ID, SRVY_YEAR, FSCSKEY, LIBNAME, ADDRESS, CITY, STABR, ZIP, CNTY, OBEREG, POPU_LSA, LEGAL_BASIS, LOCALE_CSTR, LOCALE_NAME, CENTLIB, BRANLIB, BKMOB, LIBRARIA, LOCGVT, STGVT, FEDGVT, OTHINCM, STAFFEXP, PRMATEXP, ELMATEXP, OTHMATEX, TOTEXPCO, OTHOPEXP, TOTOPEXP, BKVOL, EBOOK, AUDIO_PH, AUDIO_DL, VIDEO_PH, VIDEO_DL, ELECCOLL, VISITS, REGBOR, TOTCIR, KIDCIRCL, ELMATCIR, PHYSCIR, TOTCOLL, LOANTO, LOANFM, TOTPRO, KIDPRO, YAPRO, TOTATTEN, KIDATTEN, YAATTEN, GPTERMS, PITUSR, WIFISESS, LONGITUD, LATITUDE)
```

Examine Data Structure:
```{r}
# Examine data structure of PLS
str(PLS_FINAL)
```

Additional tidying notes:
- For new variables, change NA's to 0 where applicable
- Add variable for total locations: `CENTLIB` + `BRANLIB` + `BKMOB`
- Add variables showing the proportion of each operating revenue type
- Add variables showing the proportion of each collection expenditure type
- Create revenue and/or expenditure categories to group by expenditure level (ie: small, medium, large)

Add new variables:
```{r}
# Add Ebook proportion variable-----------------------------------------------------------------------------
PLS_FINAL <- mutate(PLS_FINAL,
                    TOT_BOOK = EBOOK + BKVOL,
                    EBOOK_PROP = EBOOK / TOT_BOOK,
                    PHBOOK_PROP = BKVOL / TOT_BOOK)
# Add per capita variables----------------------------------------------------------------------------------
PLS_FINAL <- mutate(PLS_FINAL,
                    PC_TOTCOLL = TOTCOLL / POPU_LSA,
                    )
# PLS_TOT <- mutate(PLS_TOT,
#                   TOTLOC = CENTLIB + BRANLIB + BKMOB,
#                   P_CENTLIB = CENTLIB / TOTLOC,
#                   P_BRANLIB = BRANLIB / TOTLOC,
#                   P_BKMOB = BKMOB / TOTLOC,
#                   P_PRMATEXP = PRMATEXP / TOTEXPCO,
#                   P_ELMATEXP = ELMATEXP / TOTEXPCO,
#                   P_OTHMATEX = OTHMATEX / TOTEXPCO,
#                   P_STAFFEXP = STAFFEXP / TOTOPEXP,
#                   P_TOTEXPCO = TOTEXPCO / TOTOPEXP,
#                   P_OTHOPEXP = OTHOPEXP / TOTOPEXP
#                   )
```

Make library names consistent:
Library names are inconsistent across survey years. Populate the `LIBNAME` field with the name from the 2017 survey year for all survey years.

```{r}
# Make library names consistent with most recent survey year libname
# Create new df that only includes FSCSCKEY and LIBNAME for the 2017 survey year
current_libnames <- select(PLS2017, FSCSKEY, LIBNAME)
# Change LIBNAME to proper case
current_libnames <- mutate(current_libnames, LIBNAME = str_to_title(LIBNAME))
# Replace LIBNAME with names in current_libnames
PLS_TABLEAU <- merge(current_libnames, PLS_FINAL, by = c("FSCSKEY"), all.x = TRUE)
# Change LIBNAME.y to proper case
PLS_TABLEAU <- mutate(PLS_TABLEAU, LIBNAME.y = str_to_title(LIBNAME.y))
```

## First 10 Rows

# Univariate Non-Graphical Analysis

# Univariate Graphical Analysis

# Multivariate Non-Graphical Analysis

# Multivariate Graphical Analysis

# Detailed EDA

# Statistical Tests

# Regression Analysis

# Findings

# CSV for Dashboard Creation

```{r}
# Save PLS and PLS_final to csv for use in Tableau and other dashboard creation software
write.csv(PLS_TABLEAU, file = "Public_Library_Survey_08_to_17.csv")
```

